---
title: "Programming Background"
author: "Damon D'Ambrosio"
date: "2023-06-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path="../images/")
```

## Response to prompt

I had a little exposure to R through grad school, but mostly coded in SAS. I was largely self-taught in both of these languages. Regarding R, I always felt it was more "fragmented," and hard to keep things organized. After some proper education in R, I see that this feeling was largely an artifact of poor technique on my part (e.g., poor project management, poor naming conventions for objects, not knowing how to write functions, not knowing the "theory" behind packages and functions). R is a very nice programming language. Now that I am more knowledgable on good technique, I find working with R to be much more enjoyable. R, especially with `tidyverse`, seems a lot more straightforward when it comes to data handling compared to SAS. One advantage I felt SAS had was reconstructing a prior analysis; I found it easier to read through existing SAS code to figure out what was done in the past. After learning about RMarkdown, I see that R can offer comparable (arguably superior) functionality in this regard. While I have hit a few snags with my learning here and there, I do not find R to be an overall difficult language to learn.

## R Markdown Output
```{r markdown_output}
plot(USArrests)

```
